### 2022-09-29 main/chapter_theory.tex
add some on simplified models

Electromagnetic and weak forces are unified into the electroweak
interactions with four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
$SU\!(2)_L$ and $U\!(1)$ symmetries respectively.
These would be massless if the symmetry were respected, but it is
`spontaneously' broken with the Higgs mechanism~\cite{
higgs1964broken,
englert1964broken
}%
\footnote{%
Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
Peter~Higgs has reportedly suggested ``ABEGHHKtH mechanism'' to fully credit
associated work~\cite{close2011infinity}.%
}
which adds four scalar fields (as a complex doublet) with a potential that
moves the ground state to a gauge-\emph{asymmetric} configuration.
Around this new ground state, linear combinations of the electroweak fields
acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
four scalar fields to yield the three weak bosons and the photon:
$W^1$ and $W^2$ mix into $W^+$ and $W^-$, while $W^3$ and $B$ mix into $Z$ and
$\gamma$.
The fourth scalar is the Higgs boson $h$~\cite{
glashow1959renorm,
weinberg1967model,
salam1959weak,
rubakov1999classical,
cottingham2007greenwood
}.

@@

Justified constraints on the model could have led to better predictions of
past data and might extrapolate to new phenomena through their deeper
understanding of nature.
Attempts to relate empirical parameters~\cite{
sato1979ratio,
beg1982gauge,
koide1983fermion,
denterria2012gaussian
}
or to derive mathematical structures~\cite{
goyal2010origin,
skilling2021abc,
axioms1010038
}
are curious, but will remain speculative until they provide novel benefits.

@@

Current observations do limit the masses and mixings of this new
school of MSSM particles, but their values remain very uncertain and in need
of new experimental constraints.

@@

@@

Supersymmetric modelling successfully predicted the lightest Higgs boson
mass to be $\lesssim 130\,\eV[G]$ long before its observation at
$125\,\eV[G]$~\cite{
espinosa1992higgs,
espinosa1993upper,
quiros1997constraints,
wells2020discovery,
atlas2012higgs,
atlas2012combined,
cms2012higgs
}.
This prediction, however, assumed a mass scale below $1\,\eV[T]$,
and observations at the LHC are now limiting sparticle masses
into super-$\eV[T]$ territory (in specific models).
Such limits are also troublesome for the hierarchy problem, because large
superpartner masses introduce fine-tuning problems of their
own~\cite{martin2016primer}.
Despite its current lack of observational evidence, however,
supersymmetry remains a tool for building theoretically plausible
models, and might help us to understand future discoveries.

>>>>

Electromagnetic and weak forces are unified into the electroweak
interactions with four bosons $W^1$, $W^2$, $W^3$, and $B$ from the
$SU\!(2)_L$ and $U\!(1)$ symmetries respectively.
These would be massless if the symmetry were respected, but it is
`spontaneously' broken with the Higgs mechanism~\cite{
higgs1964broken,
englert1964broken
}%
\footnote{%
Historical accuracy prefers the name `Brout–Englert–Higgs mechanism', but
Peter~Higgs has reportedly suggested ``ABEGHHKtH mechanism'' to fully credit
associated work~\cite{close2011infinity}.%
}
which a complex doublet (equivalently four scalar fields) with an associated
potential.
Minima of the potential lie in a gauge-symmetric set, and the choice made by
a system as it settles into one particular ground state spontaneously breaks
the symmetry.
Around the chosen ground state, linear combinations of the electroweak fields
acquire masses while ``eating up''~\cite{rubakov1999classical} three of the
four scalar fields to yield the three weak bosons and the photon:
$W^1$ and $W^2$ mix into $W^+$ and $W^-$, while $W^3$ and $B$ mix into $Z$ and
$\gamma$.
The fourth scalar is the Higgs boson $h$~\cite{
glashow1959renorm,
weinberg1967model,
salam1959weak,
rubakov1999classical,
cottingham2007greenwood
}.

@@

Justified constraints on the model could have led to better predictions of
past data and might extrapolate to new phenomena through their deeper
understanding of nature.
Attempts to relate the empirical parameters~\cite{
sato1979ratio,
beg1982gauge,
koide1983fermion,
denterria2012gaussian
}
or to derive the mathematical structures~\cite{
goyal2010origin,
skilling2021abc,
axioms1010038
}
are curious, but will remain only curiosities unless they provide novel
benefits.

@@

@@

Current observations do limit the masses and mixings of this new
school of MSSM particles, but their values remain very uncertain and in need
of new experimental constraints.
Diverse observable effects can be induced by parameter choices,
so we use simplified models to coarsely characterize interesting selections
of plausible effects.
Such simplified models feature only a few free variables, and decouple other
effects by, for example, sending masses to extremely large values or assuming
no mixing~\cite{
atlas2022searches,
alwall2009simplified,
alves20212simplified
}.

@@

Supersymmetric modelling successfully predicted the lightest Higgs boson
mass to be $\lesssim 130\,\eV[G]$ long before its observation at
$125\,\eV[G]$~\cite{
espinosa1992higgs,
espinosa1993upper,
quiros1997constraints,
wells2020discovery,
atlas2012higgs,
atlas2012combined,
cms2012higgs
}.
This prediction, however, assumed a mass scale below $1\,\eV[T]$,
and observations at the LHC are now limiting sparticle masses
into super-$\eV[T]$ territory (in specific models).
Such limits are also troublesome for the hierarchy problem, because large
superpartner masses introduce fine-tuning problems of their
own~\cite{martin2016primer}.
Despite its current lack of observational evidence, however,
supersymmetry remains a tool for building theoretically plausible
models, and might help us to understand future discoveries,
which could perhaps be made in \atlas\ data.
